---
title: "COVID-19 Global Analysis Using Johns Hopkins Dataset"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 4
)
library(tidyverse)
library(lubridate)
library(countrycode)
library(forecast)
library(scales)
```

## Research Question

**How has the spread and fatality rate of COVID-19 evolved across continents, and what can we forecast about its global trend?**

## Data Source and Description

- **Source**: [Johns Hopkins CSSE COVID-19 Dataset](https://github.com/CSSEGISandData/COVID-19)
- **Files Used**:
  - `time_series_covid19_confirmed_global.csv`
  - `time_series_covid19_deaths_global.csv`
- **Description**: This dataset contains daily cumulative counts of confirmed COVID-19 cases and deaths by country, beginning in January 2020.

## Data Import and Preparation

```{r data-import}
url_confirmed <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
url_deaths <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"

confirmed_raw <- read_csv(url_confirmed)
deaths_raw <- read_csv(url_deaths)

confirmed <- confirmed_raw %>%
  pivot_longer(cols = matches("^\\d"), names_to = "Date", values_to = "Confirmed") %>%
  mutate(Date = mdy(Date))

deaths <- deaths_raw %>%
  pivot_longer(cols = matches("^\\d"), names_to = "Date", values_to = "Deaths") %>%
  mutate(Date = mdy(Date))

covid_data <- confirmed %>%
  select(-Lat, -Long) %>%
  left_join(deaths %>% select(-Lat, -Long),
            by = c("Province/State", "Country/Region", "Date")) %>%
  group_by(`Country/Region`, Date) %>%
  summarise(
    Confirmed = sum(Confirmed, na.rm = TRUE),
    Deaths = sum(Deaths, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Continent = countrycode(`Country/Region`, "country.name", "continent"))
```

## Exploratory Data Analysis

### Global Cumulative Cases and Deaths

```{r global-trends}
covid_global <- covid_data %>%
  group_by(Date) %>%
  summarise(Confirmed = sum(Confirmed), Deaths = sum(Deaths), .groups = "drop")

ggplot(covid_global, aes(x = Date)) +
  geom_line(aes(y = Confirmed), color = "steelblue") +
  geom_line(aes(y = Deaths), color = "firebrick") +
  labs(
    title = "Cumulative Global COVID-19 Cases and Deaths",
    x = "Date",
    y = "Number of People",
    caption = "Data: Johns Hopkins CSSE"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()
```

### Case Fatality Rate (CFR) by Continent

```{r cfr-continent}
cfr_data <- covid_data %>%
  filter(!is.na(Continent)) %>%
  group_by(Date, Continent) %>%
  summarise(Confirmed = sum(Confirmed), Deaths = sum(Deaths), .groups = "drop") %>%
  mutate(CFR = Deaths / Confirmed)

ggplot(cfr_data, aes(x = Date, y = CFR, color = Continent)) +
  geom_line() +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "COVID-19 Case Fatality Rate by Continent",
    x = "Date",
    y = "CFR (%)",
    caption = "CFR = Deaths / Confirmed Cases"
  ) +
  theme_minimal()
```

## Modeling and Forecasting

### 30-Day Forecast Using ARIMA

```{r arima-forecast}
ts_confirmed <- ts(covid_global$Confirmed, frequency = 7)
fit <- auto.arima(ts_confirmed)
fc <- forecast(fit, h = 30)

autoplot(fc) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Forecast of Global Confirmed COVID-19 Cases (Next 30 Days)",
    x = "Time (in Weeks)",
    y = "Cumulative Confirmed Cases",
    caption = "ARIMA forecast based on past cumulative data"
  ) +
  theme_minimal()
```

## Limitations and Biases

- **Underreporting**: Some countries may lack resources for widespread testing.
- **Reporting Lag**: Case and death numbers may not be reported in real time.
- **Data Gaps**: Incomplete or inconsistent data for certain regions.
- **Forecasting Model Limitations**: ARIMA assumes historical patterns continue, which may not hold under new policy changes or variants.

## Conclusion

This analysis demonstrates temporal trends and differences in case fatality across continents. While ARIMA provides a basic predictive tool, real-world events like new variants and public health interventions must be considered to interpret the results accurately.

## References

1. Johns Hopkins University CSSE. (2025). COVID-19 Data Repository. [GitHub](https://github.com/CSSEGISandData/COVID-19)
2. Hyndman, R.J., & Athanasopoulos, G. (2018). *Forecasting: Principles and Practice*. OTexts.
